---
title: "Dust products"
author: "Athanasios Natsis"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  slidy_presentation: default
  beamer_presentation:
    latex_engine: xelatex
  ioslides_presentation:
    widescreen: true
    smaller: true
subtitle: Some options
---

<!-- ## Create a Dust product {.smaller} -->

<!-- 1. Polar Multi-sensor Aerosol product (PMAp) -->


## Polar Multi-sensor Aerosol product (PMAp) {.smaller}
Climate record 2007--2019 (PMAp CDR)

Output

- AOD (GOME-2)
- Aersool types and clouds (IASI, AVHRR)
- Dust as non operational product...
- PMAp (dust-dominated total AOD contribution from other aerosol types than dust)
- LEO

Aerosol types:
 ocean fine mode, ocean coarse mode, thick biomass burning, desert dust, volcanic ash/thick dust, volcanic ash with SO2, aerosol contaminated cloud, ash contaminated cloud, unclassified


## Polar Multi-sensor Aerosol product (PMAp) {.smaller}
Inputs:

- GOME-2 Fundamental Data Record (FDR) Level 1C
- IASI FDR Level-1C (NRT)
- AVHRR Level 1B NRT


## Polar Multi-sensor Aerosol product (PMAp) {.smaller}
Algorithm:

- cloud identification and correction using collocated AVHRR observations
- discrimination between aerosols and clouds 
- preliminary classification of aerosols, including desert dust detection.
- AODs also for partly cloudy footprints, 
<!-- if the cloud fraction is not too large and reflectances can be corrected for cloud contamination -->

- retrieval of estimated AOD using unpolarised GOME-2 reflectances and a Look-Up-Table
- set of AOD estimates is retrieved for each candidate aerosol model.

- find the best fit between the measured and modelled reflectances and Stokes
fractions

- best model (LUTs from  9 models), for each aerosol, aerosol type is set and final AOD from the AODs retrieved
at the second step is selected

<!-- - 4 aerosol models with different absorption and vertical profile, used in the forward model represent dust -->

Validation:

- AERONET



## Polar Multi-sensor Aerosol product (PMAp) {.smaller}

PMAp --> AODs @ 550nm

- not possible to define the dust fraction (or AOD related to dust particles only) from a PMAp total AOD
- Define the dust-dominated total AOD
where the aerosol class is set to “desert dust” or "thick dust" 
for pixels which the assumed dominant aerosol type is dust

“total AOD” 
“dust-dominated total AOD” 

WORKS  but:

- for subset of data/domain
- against specially filtered AERONET data
- better describing dust occurrence and variation
- sensitive enough for high AOD cases involving dust
- PMAp observations cannot be used to define DAOD global variation




## Thanos AOD retrieval (we need a nice name)
Inputs:

- Temperature @10.8μm        (MSG --> update σε MTG)
- Land Surface Temperature   (ERA5 --> update σε MTG[^1])
- Humidity                   (ERA5)
- Constant from MODTRAN4     (more model runs)
- Cloud mask                 (MSG/MTG)

[^1]: pre-operational

## Thanos AOD retrieval (we need a nice name)
Difficulties:

- Discrimination of dust from clouds
- Define/choose a rolling pristine sky
- ERA5 latency 5 days for preliminary, 2-3 months for final

Validation:

- AERONET
- Sentinel-3?

Upgrades:

- Automation
- Machine learning

<!--

## Data Visualization {.flexbox .vcenter}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
```


```{r}
ggplot(cars, aes(x = speed, y = dist)) +
  geom_point(color = "steelblue", size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", color = "darkred") +
  labs(title = "Car Stopping Distance vs Speed",
       x = "Speed (mph)",
       y = "Stopping Distance (ft)") +
  theme_minimal()
```

## Data visualization {.flexbox .vcenter}

```{r}
car_stats <- cars %>%
  summarize(
    Mean_Speed = mean(speed),
    Mean_Distance = mean(dist),
    Correlation = cor(speed, dist)
  )

knitr::kable(car_stats, caption = "Summary Statistics")
```
-->
